Cross-Validation Approaches for Penalized Cox Regression

### 2024-01-11 main.tex
Finished revisions based on referee feedback

In this paper, we propose an alternative way to carry out cross validation for Cox regression. Instead of applying cross-validation to the partial likelihood directly, we propose applying cross validation to the linear predictors of the regression model. Through simulation studies, we compare these proposed methods with existing cross-validation approaches for LASSO penalized Cox regression in both low- and high-dimensional settings. We find that the linear predictor approach avoids the conservatism of the deviance residual and \citet{Verweij1993} approaches as well as the numerical instability of ordinary cross-validation. The proposed approach offers the best combination of performance and stability, and we recommend using it for regularization parameter selection in penalized Cox models.  We conclude by applying both proposed and existing approaches to a high-dimensional study of gene expression and survival in lung cancer patients.

@@

In this paper, we focus upon the LASSO penalty $P_{\lambda}(\bbeta) = \lambda \sum_{j} |\beta_{j}|$ as it is most common, although the need to select regularization parameters arises in all penalized regression models, such as elastic net \citep{Zou2005}, smoothly clipped absolute deviations (SCAD) penalty \citep{FanLi2001}, and minimax concave penalty (MCP) \citep{Zhang2010}. Indeed, cross-validation can be used for model selection in Cox regression even without the presence of a penalty, so the relevance of this question is wide-ranging. LASSO-penalized Cox regression is particularly useful for high dimensional data where the number of covariates $p > n$, such as predicting overall survival in cancer patients based on genome-wide expression measurements. LASSO estimates yield a sparse model where some coefficients are exactly zero. This sparsity pattern changes as we vary $\lam$: at large values of $\lambda$, most or all of the coefficients are 0, but as $\lam$ decreases, more covariates are selected. Selecting $\lambda$ is critical to LASSO estimation in the sense that the model's accuracy suffers if $\lam$ is either too large or too small.

@@

\subsection{Cross validated partial likelihood} 
\label{Sec:cox-cv-existing}

@@

\begin{figure}
  \centering
  \includegraphics[height= 12cm ]{figure_1_new.png}
  \caption{\label{Fig:methods} Illustration of $K$-fold cross validation methods for Cox regression when $k = 1$}
\end{figure}

%%%%%%%%%%%%%%%%%%%%
\subsection{Cross validated deviance residuals}
%%%%%%%%%%%%%%%%%%%%

The fundamental challenge of conducting cross validation for Cox regression is that the baseline hazard is not estimated from the model, which prevents us from calculating prediction errors (residuals).  However, if one takes additional steps to estimate the baseline hazard in addition to estimating the regression coefficients, then such residuals can be calculated.  The deviance residual, a normalized form of the Martingale residual \citep{Therneau1990}, is a natural candidate for quantifying prediction error to use in quantifying the cross-validation error. Originally proposed for model diagnostics, deviance residuals have also been used as a predictive accuracy measure, particularly in the random survival forest literature \citep{LeBlanc1992, Ishwaran2004}. In the Cox model, the Martingale residuals $\{m_i\}_{i=1}^n$ and deviance residuals $\{d_i\}_{i=1}^n$ are given by
\begin{align}
  \label{eq:res-m}
  m_{i} &= \delta_{i} - H_{0}(t_{i})\exp(\x_i\Tr\bbeta), \\
  \label{eq:res-d}
  d_{i} &= \text{sign}(m_{i})\sqrt{-2(m_i + \delta_{i}\log(\delta_{i} - m_i))}.  
\end{align}
where $H_0$ is the cumulative baseline hazard.

@@

The sum of squared cross-validated deviance residuals, $\sum_{i}\hat{d}_{i}^2$, are then used as the cross validated error, analogous to using the residual sum of squares in linear regression. We refer to this method as the \emph{cross-validated deviance residuals} approach.

%%%%%%%%%%%%%%%%%%%%
\subsection{Cross validated linear predictors}
\label{Sec:linear-predictor}
%%%%%%%%%%%%%%%%%%%%

@@

Simulation studies were conducted to compare how the methods described in Section~\ref{Sec:methods} behave relative to each other in selecting the tuning parameter $\lambda$ and the corresponding model for LASSO penalized regression. For these simulation studies, the covariate matrix $X$, consisting of $n$ observations and $p$ covariates, was generated from independent $\text{Normal}(0, 1)$ samples. Given $\x_i$ and a coefficient vector $\bbeta$, survival times were generated from an exponential distribution with rate parameter $\exp(\x_i \Tr \bbeta)$. Censoring status indicators were drawn independently from the binomial distribution. All simulations were implemented in R \citep{R}.

@@

Estimation accuracy was measured using the mean squared error of the coefficients ($\MSE$). We define $\bbh(\lambda)$ as the vector of coefficients estimated by a Cox model with LASSO penalty at $\lambda$, and define $\bbeta^*$ as the true value of $\bbeta$ (the value used to generate the data). Mean squared error is defined to be 
\begin{equation}
\MSE = E\| \bbh(\lambda) - \bbeta^* \| ^2.
\end{equation}

@@

We also measure the predictive accuracy via the mean squared prediction error, or Brier score \citep{VanHouwelingen2011}. Let $\hat{S}(t_0|\lambda, \x)$ denote the predicted probability for an individual with covariates $\x$ to survive beyond time $t_0$, and let $y(t_0) = \mathbb{1}\left\{ t > t_{0}\right\}$ indicate whether the individual actually survived beyond time $t_0$. Then the Brier score is defined to be
\begin{equation}
\text{Brier}(y(t_0), \hat{S}(t_0|\lambda, \x)) = \{ y(t_0) - \hat{S}(t_0|\lambda, \x) \}^2.
\end{equation}
In our simulation study, for each data set, we generated another independent data set with 1000 uncensored observations. We computed Brier scores for all of the cross-validation methods for all individuals in this test set, setting $t_0$ to the median survival time.

@@

The final metric we included in our study was Harrell's C Index \citep{HarrellJr1984}. The C Index is a rank-based statistic that measures the concordance between the linear predictor of the selected model and the observed outcome. Suppose we arbitrarily select two individuals; if the individual with the higher predicted risk also died first, then this would be considered as a concordant pair. The C Index considers all possible such pairs and estimates of proportion of those pairs that are concordant. A C Index of 0.5 means the prediction is no better than flipping a coin and C Index of 1 means perfect concordance. We computed the C Index using the independent test set of 1000 individuals described above.

@@

\caption{\label{Tab:sim} Predictive accuracy of four cross-validation methods under scenarios with varying dimension and signal strength (standard deviations given in parentheses).}
\centering
\begin{tabular}[t]{lccccc}
\toprule
 Method & $\lambda$ & log(MSE Ratio) &Brier Score & Kullback Leibler & C Index\\
\midrule
&\multicolumn{5}{c}{\textbf{Scenario 1}: n = 150, p = 50, 5 non-zero features, weak signal}\\
\cline{2-6}
V\&VH  & 0.098 (0.030) & 1.561 (0.851) & 0.245 (0.014) & 0.685 (0.030) & 0.610 (0.035) \\
Basic  & 0.088 (0.025) & 1.531 (0.838) & 0.244 (0.013) & 0.683 (0.029) & 0.614 (0.029) \\
LinearPred  & 0.092 (0.029) & 1.543 (0.846) & 0.245 (0.014) & 0.684 (0.030) & 0.612 (0.034) \\
DevResid  & 0.098 (0.029) & 1.547 (0.848) & 0.245 (0.014) & 0.684 (0.030) & 0.611 (0.035) \\
\addlinespace
&\multicolumn{5}{c}{\textbf{Scenario 2}: n = 150, p = 50, 5 non-zero features, strong signal}\\
\cline{2-6}
V\&VH  & 0.071 (0.010) & 1.586 (0.989) & 0.176 (0.011) & 0.526 (0.027) & 0.750 (0.012) \\
Basic  & 0.066 (0.011) & 1.544 (0.971) & 0.176 (0.011) & 0.526 (0.028) & 0.749 (0.013) \\
LinearPred  & 0.067 (0.010) & 1.540 (0.981) & 0.176 (0.011) & 0.525 (0.028) & 0.750 (0.012) \\
DevResid  & 0.094 (0.011) & 1.995 (0.841) & 0.181 (0.011) & 0.539 (0.026) & 0.751 (0.013) \\
\addlinespace
&\multicolumn{5}{c}{\textbf{Scenario 3}: n = 400, p = 10000, 20 non-zero features, weak signal}\\
\cline{2-6}
V\&VH  & 0.126 (0.019) & 2.958 (0.416) & 0.251 (0.012) & 0.696 (0.026) & 0.589 (0.044) \\
Basic  & 0.103 (0.022) & 2.880 (0.437) & 0.244 (0.015) & 0.680 (0.031) & 0.603 (0.039) \\
LinearPred  & 0.105 (0.023) & 2.886 (0.437) & 0.244 (0.015) & 0.681 (0.031) & 0.602 (0.039) \\
DevResid  & 0.107 (0.022) & 2.895 (0.433) & 0.245 (0.014) & 0.683 (0.03) & 0.602 (0.039) \\
\addlinespace
&\multicolumn{5}{c}{\textbf{Scenario 4}: n = 400, p = 10000, 20 non-zero features, strong signal}\\
\cline{2-6}
V\&VH  & 0.084 (0.012) & 3.799 (0.548) & 0.176 (0.03) & 0.528 (0.069) & 0.763 (0.048) \\
Basic  & 0.059 (0.007) & 3.455 (0.578) & 0.154 (0.019) & 0.468 (0.049) & 0.780 (0.031) \\
LinearPred  & 0.060 (0.007) & 3.479 (0.574) & 0.155 (0.019) & 0.470 (0.049) & 0.780 (0.030) \\
DevResid  & 0.076 (0.018) & 3.708 (0.518) & 0.166 (0.024) & 0.502 (0.057) & 0.772 (0.049) \\
\bottomrule
\end{tabular}
\end{table}

@@

\begin{table}[!htb]
\setlength{\tabcolsep}{3pt}
\caption{\label{Tab:power} Number of non-zero features (true positives) selected by each cross-validation method, under scenarios with varying dimension and signal strength. For weak and strong signal strength, nonzero coefficients are set to 0.3 and 0.6 respectively.}
\centering
\begin{tabular}[t]{l cc c cc c cc}
\toprule
&\multicolumn{2}{c}{n = 150, p = 50} & & \multicolumn{2}{c}{n = 300, p = 1000} & & \multicolumn{2}{c}{n = 400, p = 10000} \\
&\multicolumn{2}{c}{$\#$ non-zero features: 5} & & \multicolumn{2}{c}{$\#$ non-zero features: 20} & & \multicolumn{2}{c}{$\#$ non-zero features: 20} \\
\cline{2-3} \cline{5-6} \cline{8-9}
& weak signal & strong signal & & weak signal & strong signal && weak signal & strong signal\\
V\&VH        & 3.87 & 5& & 10.96 & 19.61 & & 6.67 & 18.38\\
Basic          & 4.06 & 5& & 13.68 & 19.96 & & 9.58 & 19.87\\
LinearPred  & 3.99 & 5& & 13.46 & 19.95 & & 9.40 & 19.85\\
DevResid    & 3.90 & 5& & 12.96 & 19.38 & & 8.98 & 19.05\\
\bottomrule
\end{tabular}
\end{table}

@@

\begin{figure}[!htb]
  \centering
  \includegraphics[width=\textwidth]{figure_3.png}
  \caption{\label{Fig:stability} Simulations illustrating the stability issues with the basic approach. For the figure on the left, $n = 100$, fold assignments are unbalanced, and the vertical axis is the proportion of undefined partial likelihoods. For the figure on the right, $n = 120$, fold assignments are balanced so that the number of events in each fold is the same, and the vertical axis is the standard deviation of the squared error loss. Basic cross-validation is much more variable than other approaches when the number of events in each fold is small.}
\end{figure}

@@

\subsection {Leave-One-Out Cross Validation}
\label{Sec:loocv}

As discussed earlier, basic cross-validation cannot be used to carry out leave-one-out cross validation (LOOCV) for Cox regression models. The other three approaches (V\&VH, linear predictor, and deviance residual), however, are not subject to the stability issues described in Section~\ref{Sec:stability}, and can be used to implement LOOCV.  Here, we compare these three approaches for carrying out LOOCV, and also compare LOOCV to 10-fold CV.

\begin{figure}[!htb]
  \centering
  \includegraphics[width=\textwidth]{figure_4.png}
  \caption{\label{Fig:loocv}Simulation comparing LOO (Leave-One-Out) and 10-fold CV across three approaches (V\&VH, linear predictor, and deviance residual). With $N = 200$ replications, the data sets generated have dimension $n = 120$ and $p = 1000$. The vertical axis is the $\log$(MSE ratio). Each panel represents one approach. In each panel, the LOO version and 10-fold version of the same approach are plotted. In all three panels, the 10-fold basic cross-validation is included as a baseline for comparison (green line).}
\end{figure}

@@

All of the cross-validation approaches described in Section~\ref{Sec:methods} use the Cox partial likelihood, in one way or another, as the criterion upon which model selection is based.  In this section, we compare these likelihood-based approaches against an alternative approach of cross validation using concordance, also known as the cross-validated C Index.  This is a widely used approach for conducting cross-validation with time-to-event outcomes, particularly for machine learning models \citep{Subramanian2011,Simon2011a}, and was recently added as an option in the \texttt{glmnet} package (\texttt{type.measure="C"}). This approach typically proceeds by constructing cross-validated linear predictors, as in Section~\ref{Sec:linear-predictor}, but then using those predictors to calculate concordance as opposed to partial likelihood.

@@

\begin{figure}[!htb]
  \centering
  \includegraphics[height= 15cm ]{figure_auc.png}
  \caption{\label{Fig:cvauc}Comparing Partial Likelihood and C Index both built from Cross-validated Linear Predictors. The horizontal axis represents the strength of signal $s$ in the data. The left figures are under the low dimension scenarios where p = 50 and n = 150. The right figures are under the high dimension scenarios where p = 1000 and n = 120. The expected percentages of censoring for both scenarios are 30 $\%$. The top row are box plots of the selected tuning parameter $\lambda$. The mid row plots the $\log(\text{relative Squared Error Loss})$, with respect to the oracle model. The bottom row plots the Brier Scores. }
\end{figure}

@@

\section{Application to Real Data}

@@

\begin{table}[ht]
\centering
\caption{\label{tb:shedden} The C Index and Brier Score of all elastic net models selected by cross-validation methods }
\begin{tabular}{lccccccc}
\hline
  & \multicolumn{7}{c}{Cross-validated C Index} \\ 
  & $\alpha = 0.1$ & $\alpha = 0.2$ & $\alpha = 0.3$ & $\alpha = 0.5$ & $\alpha = 0.7$ & $\alpha = 0.9$ & $\alpha = 1$ \\ \cline{2-8}
V \& VH        & 0.645 &0.643 &0.630 &0.626 &0.617 &0.617 &0.617 \\
Basic          & 0.649 &0.650 &0.650 &0.649 &0.650 &0.651 &0.651 \\
Linear Pred    & 0.648 &0.650 &0.650 &0.649 &0.648 &0.617 &0.617 \\
Deviance Resid & 0.645 &0.647 &0.646 &0.626 &0.617 &0.617 &0.617 \\ 
\addlinespace
  & \multicolumn{7}{c}{Brier Score} \\ 
  & $\alpha = 0.1$ & $\alpha = 0.2$ & $\alpha = 0.3$ & $\alpha = 0.5$ & $\alpha = 0.7$ & $\alpha = 0.9$ & $\alpha = 1$ \\ \cline{2-8}
V \& VH        &0.231 &0.232 &0.243 &0.245 &0.249 &0.249 &0.249 \\
Basic          &0.226 &0.226 &0.225 &0.224 &0.221 &0.221 &0.221 \\
Linear Pred    &0.228 &0.226 &0.225 &0.224 &0.224 &0.249 &0.249 \\
Deviance Resid &0.231 &0.229 &0.228 &0.245 &0.249 &0.249 &0.249 \\ 
\hline
\end{tabular}
\end{table}

@@

\begin{figure}[!htb]
  \centering
  \includegraphics[height= 8cm ]{shedden.png}
  \caption{\label{fig:CVE}Rescaled Cross Validated Errors at $\alpha = 0.3$. The horizonal axis is $\log(\lambda)$. The vertical axis is the cross-validated error rescaled to a range from 0 to 1 in order to allow plotting of the four methods on a common axis. The vertical dashed lines represent the $\lambda$ selected by a specific method. The linear predictor approach selected the smallest $\lambda$ and Verweij and Van Houwelingen's approach selected the greatest $\lambda$.}
\end{figure}

@@

\begin{table}[ht]
\centering
\caption{\label{Tab:go} Five most significantly enriched molecular functions among genes selected by the linear predictor method with the elastic net ($\alpha=0.3$).  The five terms have a false discovery rate of 3\% using the Benjamini-Hochberg method. Enrichment: frequency of the annotation among the selected genes divided by its frequency in the genome as a whole.}
\begin{tabular}{llrrr}
\toprule
ID & Name & Selected & Enrichment & $p$-value \\
\midrule
GO:0008134 & transcription factor binding & 10 & 4.1 & 0.00015 \\
GO:0035257 & nuclear hormone receptor binding & 5 & 7.4 & 0.00057 \\
GO:0019869 & chloride channel inhibitor activity & 2 & 53.2 & 0.00061 \\
GO:0003700 & DNA-binding transcription factor activity & 14 & 2.6 & 0.00068 \\
GO:0005102 & signaling receptor binding & 16 & 2.4 & 0.00083 \\
\bottomrule
\end{tabular}
\end{table}

@@

Although we focused upon the penalized Cox regression with a LASSO penalty in this work, the methods that we analyzed are applicable in any context that involves conducting model selection for survival models that are built upon the Cox partial likelihood, such as fitting penalized Cox regression with other penalty functions (elastic net \citep{Zou2005}, SCAD \citep{FanLi2001}, MCP \citep{Zhang2010}, precision lasso \citep{Wang2018}), other penalty structures (group selection, bi-level selection), and selecting functional forms of covariates in unpenalized Cox regression.

>>>>

In this paper, we propose an alternative way to carry out cross-validation for Cox regression. Instead of applying cross-validation to the partial likelihood directly, we propose applying cross-validation to the linear predictors of the regression model. Through simulation studies, we compare these proposed methods with existing cross-validation approaches for lasso-penalized Cox regression in both low- and high-dimensional settings. We find that the linear predictor approach avoids the conservatism of the deviance residual and \citet{Verweij1993} approaches as well as the numerical instability of ordinary cross-validation. The proposed approach offers the best combination of performance and stability, and we recommend using it for regularization parameter selection in penalized Cox models.  We conclude by applying both proposed and existing approaches to a high-dimensional study of gene expression and survival in lung cancer patients.

@@

In this paper, we focus upon the lasso penalty $P_{\lambda}(\bbeta) = \lambda \sum_{j} |\beta_{j}|$ as it is most common, although the need to select regularization parameters arises in all penalized regression models, such as elastic net \citep{Zou2005}, smoothly clipped absolute deviations (SCAD) penalty \citep{FanLi2001}, and minimax concave penalty (MCP) \citep{Zhang2010}. Indeed, cross-validation can be used for model selection in Cox regression even without the presence of a penalty, so the relevance of this question is wide-ranging. Lasso-penalized Cox regression is particularly useful for high dimensional data where the number of covariates $p > n$, such as predicting overall survival in cancer patients based on genome-wide expression measurements. Lasso estimates yield a sparse model where some coefficients are exactly zero. This sparsity pattern changes as we vary $\lam$: at large values of $\lambda$, most or all of the coefficients are 0, but as $\lam$ decreases, more covariates are selected. Selecting $\lambda$ is critical to lasso estimation in the sense that the model's accuracy suffers if $\lam$ is either too large or too small.

@@

\subsection{Cross-validated partial likelihood} 
\label{Sec:cox-cv-existing}

@@

\begin{figure}
  \centering
  \includegraphics[height= 12cm ]{figure_1_new.png}
  \caption{\label{Fig:methods} Illustration of $K$-fold cross-validation methods for Cox regression when $k = 1$}
\end{figure}

%%%%%%%%%%%%%%%%%%%%
\subsection{Cross-validated deviance residuals}
%%%%%%%%%%%%%%%%%%%%

The fundamental challenge of conducting cross-validation for Cox regression is that the baseline hazard is not estimated from the model, which prevents us from calculating prediction errors (residuals).  However, if one takes additional steps to estimate the baseline hazard in addition to estimating the regression coefficients, then such residuals can be calculated.  The deviance residual, a normalized form of the Martingale residual \citep{Therneau1990}, is a natural candidate for quantifying prediction error to use in quantifying the cross-validation error. Originally proposed for model diagnostics, deviance residuals have also been used as a predictive accuracy measure, particularly in the random survival forest literature \citep{LeBlanc1992, Ishwaran2004}. In the Cox model, the Martingale residuals $\{m_i\}_{i=1}^n$ and deviance residuals $\{d_i\}_{i=1}^n$ are given by
\begin{align}
  \label{eq:res-m}
  m_{i} &= \delta_{i} - H_{0}(t_{i})\exp(\x_i\Tr\bbeta), \\
  \label{eq:res-d}
  d_{i} &= \text{sign}(m_{i})\sqrt{-2(m_i + \delta_{i}\log(\delta_{i} - m_i))}.  
\end{align}
where $H_0$ is the cumulative baseline hazard.

@@

The sum of squared cross-validated deviance residuals, $\sum_{i}\hat{d}_{i}^2$, are then used as the cross-validated error, analogous to using the residual sum of squares in linear regression. We refer to this method as the \emph{cross-validated deviance residuals} approach.

%%%%%%%%%%%%%%%%%%%%
\subsection{Cross-validated linear predictors}
\label{Sec:linear-predictor}
%%%%%%%%%%%%%%%%%%%%

@@

Simulation studies were conducted to compare how the methods described in Section~\ref{Sec:methods} behave relative to each other in selecting the tuning parameter $\lambda$ and the corresponding model for lasso penalized regression. For these simulation studies, the covariate matrix $X$, consisting of $n$ observations and $p$ covariates, was generated from independent $\text{Normal}(0, 1)$ samples. Given $\x_i$ and a coefficient vector $\bbeta$, survival times were generated from an exponential distribution with rate parameter $\exp(\x_i \Tr \bbeta)$. Censoring status indicators were drawn independently from the binomial distribution. All simulations were implemented in R \citep{R}.

@@

Estimation accuracy was measured using the mean squared error of the coefficients ($\MSE$). We define $\bbh(\lambda)$ as the vector of coefficients estimated by a Cox model with lasso penalty at $\lambda$, and define $\bbeta^*$ as the true value of $\bbeta$ (the value used to generate the data). Mean squared error is defined to be 
\begin{equation}
\MSE = E\| \bbh(\lambda) - \bbeta^* \| ^2.
\end{equation}

@@

We also measure the predictive accuracy via the mean squared prediction error, or Brier score \citep{VanHouwelingen2011}. Let $\hat{S}(t_0|\lambda, \x)$ denote the predicted probability for an individual with covariates $\x$ to survive beyond time $t_0$, and let $y(t_0) = \mathbb{1}\left\{ t > t_{0}\right\}$ indicate whether the individual actually survived beyond time $t_0$. Then the Brier score is defined to be
\begin{equation}
\text{Brier}(y(t_0), \hat{S}(t_0|\lambda, \x)) = \{ y(t_0) - \hat{S}(t_0|\lambda, \x) \}^2.
\end{equation}
In our simulation study, for each data set, we generated another independent data set with 1000 uncensored observations to assess prediction accuracy. We computed Brier scores for all of the cross-validation methods for all individuals in this test set, setting $t_0$ to the median survival time.

@@

The final metric we included in our study was Harrell's C Index \citep{HarrellJr1984}. The C Index is a rank-based statistic that measures the concordance between the linear predictor of the selected model and the observed outcome. Suppose we arbitrarily select two individuals; if the individual with the higher predicted risk also died first, then this would be considered as a concordant pair. The C Index considers all possible such pairs and estimates of proportion of those pairs that are concordant. A C Index of 0.5 means the prediction is no better than flipping a coin and C Index of 1 means perfect concordance. We computed the C Index using the same set of 1000 independent observations as the Brier score.

@@

\caption{\label{Tab:sim} Predictive accuracy of four cross-validation methods under scenarios with varying dimension and signal strength (standard deviations given in parentheses; Section~\ref{Sec:accuracy}.}
\centering
\begin{tabular}[t]{lccccc}
\toprule
 Method & $\lambda$ & log(MSE Ratio) &Brier Score & Kullback Leibler & C Index\\
\midrule
&\multicolumn{5}{c}{\textbf{Scenario 1}: n = 150, p = 50, 5 non-zero features, weak signal}\\
\cline{2-6}
V\&VH  & 0.098 (0.030) & 1.561 (0.851) & 0.245 (0.014) & 0.685 (0.030) & 0.610 (0.035) \\
Basic  & 0.088 (0.025) & 1.531 (0.838) & 0.244 (0.013) & 0.683 (0.029) & 0.614 (0.029) \\
LinearPred  & 0.092 (0.029) & 1.543 (0.846) & 0.245 (0.014) & 0.684 (0.030) & 0.612 (0.034) \\
DevResid  & 0.098 (0.029) & 1.547 (0.848) & 0.245 (0.014) & 0.684 (0.030) & 0.611 (0.035) \\
\addlinespace
&\multicolumn{5}{c}{\textbf{Scenario 2}: n = 150, p = 50, 5 non-zero features, strong signal}\\
\cline{2-6}
V\&VH  & 0.071 (0.010) & 1.586 (0.989) & 0.176 (0.011) & 0.526 (0.027) & 0.750 (0.012) \\
Basic  & 0.066 (0.011) & 1.544 (0.971) & 0.176 (0.011) & 0.526 (0.028) & 0.749 (0.013) \\
LinearPred  & 0.067 (0.010) & 1.540 (0.981) & 0.176 (0.011) & 0.525 (0.028) & 0.750 (0.012) \\
DevResid  & 0.094 (0.011) & 1.995 (0.841) & 0.181 (0.011) & 0.539 (0.026) & 0.751 (0.013) \\
\addlinespace
&\multicolumn{5}{c}{\textbf{Scenario 3}: n = 400, p = 10000, 20 non-zero features, weak signal}\\
\cline{2-6}
V\&VH  & 0.126 (0.019) & 2.958 (0.416) & 0.251 (0.012) & 0.696 (0.026) & 0.589 (0.044) \\
Basic  & 0.103 (0.022) & 2.880 (0.437) & 0.244 (0.015) & 0.680 (0.031) & 0.603 (0.039) \\
LinearPred  & 0.105 (0.023) & 2.886 (0.437) & 0.244 (0.015) & 0.681 (0.031) & 0.602 (0.039) \\
DevResid  & 0.107 (0.022) & 2.895 (0.433) & 0.245 (0.014) & 0.683 (0.03) & 0.602 (0.039) \\
\addlinespace
&\multicolumn{5}{c}{\textbf{Scenario 4}: n = 400, p = 10000, 20 non-zero features, strong signal}\\
\cline{2-6}
V\&VH  & 0.084 (0.012) & 3.799 (0.548) & 0.176 (0.03) & 0.528 (0.069) & 0.763 (0.048) \\
Basic  & 0.059 (0.007) & 3.455 (0.578) & 0.154 (0.019) & 0.468 (0.049) & 0.780 (0.031) \\
LinearPred  & 0.060 (0.007) & 3.479 (0.574) & 0.155 (0.019) & 0.470 (0.049) & 0.780 (0.030) \\
DevResid  & 0.076 (0.018) & 3.708 (0.518) & 0.166 (0.024) & 0.502 (0.057) & 0.772 (0.049) \\
\bottomrule
\end{tabular}
\end{table}

@@

\begin{table}[!htb]
\setlength{\tabcolsep}{3pt}
\caption{\label{Tab:power} Number of non-zero features (true positives) selected by each cross-validation method, under scenarios with varying dimension and signal strength (Section~\ref{Sec:power}). For weak and strong signal strength, nonzero coefficients are set to 0.3 and 0.6 respectively.}
\centering
\begin{tabular}[t]{l cc c cc c cc}
\toprule
&\multicolumn{2}{c}{n = 150, p = 50} & & \multicolumn{2}{c}{n = 300, p = 1000} & & \multicolumn{2}{c}{n = 400, p = 10000} \\
&\multicolumn{2}{c}{$\#$ non-zero features: 5} & & \multicolumn{2}{c}{$\#$ non-zero features: 20} & & \multicolumn{2}{c}{$\#$ non-zero features: 20} \\
\cline{2-3} \cline{5-6} \cline{8-9}
& weak signal & strong signal & & weak signal & strong signal && weak signal & strong signal\\
V\&VH        & 3.87 & 5& & 10.96 & 19.61 & & 6.67 & 18.38\\
Basic          & 4.06 & 5& & 13.68 & 19.96 & & 9.58 & 19.87\\
LinearPred  & 3.99 & 5& & 13.46 & 19.95 & & 9.40 & 19.85\\
DevResid    & 3.90 & 5& & 12.96 & 19.38 & & 8.98 & 19.05\\
\bottomrule
\end{tabular}
\end{table}

@@

\begin{figure}[!htb]
  \centering
  \includegraphics[width=\textwidth]{figure_3.png}
  \caption{\label{Fig:stability} Simulations illustrating the stability issues with the basic approach (Section~\ref{Sec:stability}). For the figure on the left, $n = 100$, fold assignments are unbalanced, and the vertical axis is the proportion of undefined partial likelihoods. For the figure on the right, $n = 120$, fold assignments are balanced so that the number of events in each fold is the same, and the vertical axis is the standard deviation of the squared error loss. Basic cross-validation is much more variable than other approaches when the number of events in each fold is small.}
\end{figure}

@@

\subsection {Leave-One-Out Cross-Validation}
\label{Sec:loocv}

As discussed earlier, basic cross-validation cannot be used to carry out leave-one-out cross-validation (LOOCV) for Cox regression models. The other three approaches (V\&VH, linear predictor, and deviance residual), however, are not subject to the stability issues described in Section~\ref{Sec:stability}, and can be used to implement LOOCV.  Here, we compare these three approaches for carrying out LOOCV, and also compare LOOCV to 10-fold CV.

\begin{figure}[!htb]
  \centering
  \includegraphics[width=\textwidth]{figure_4.png}
  \caption{\label{Fig:loocv} Simulation comparing LOO (Leave-One-Out) and 10-fold CV across three approaches (V\&VH, linear predictor, and deviance residual) as described in Section~\ref{Sec:loocv}. With $N = 200$ replications, the data sets generated have dimension $n = 120$ and $p = 1000$. The vertical axis is the $\log$(MSE ratio). Each panel represents one approach. In each panel, the LOO version and 10-fold version of the same approach are plotted. In all three panels, the 10-fold basic cross-validation is included as a baseline for comparison (green line).}
\end{figure}

@@

All of the cross-validation approaches described in Section~\ref{Sec:methods} use the Cox partial likelihood, in one way or another, as the criterion upon which model selection is based.  In this section, we compare these likelihood-based approaches against an alternative approach of cross-validation using concordance, also known as the cross-validated C Index.  This is a widely used approach for conducting cross-validation with time-to-event outcomes, particularly for machine learning models \citep{Subramanian2011,Simon2011a}, and was recently added as an option in the \texttt{glmnet} package (\texttt{type.measure="C"}). This approach typically proceeds by constructing cross-validated linear predictors, as in Section~\ref{Sec:linear-predictor}, but then using those predictors to calculate concordance as opposed to partial likelihood.

@@

\begin{figure}[!htb]
  \centering
  \includegraphics[height= 15cm ]{figure_auc.png}
  \caption{\label{Fig:cvauc} Comparing Partial Likelihood and C Index both built from cross-validated Linear Predictors (Section~\ref{Sec:CIndex}). The horizontal axis represents the strength of signal $s$ in the data. The left figures are under the low dimension scenarios where p = 50 and n = 150. The right figures are under the high dimension scenarios where p = 1000 and n = 120. The expected percentages of censoring for both scenarios are 30 $\%$. The top row are box plots of the selected tuning parameter $\lambda$. The mid row plots the $\log(\text{relative Squared Error Loss})$, with respect to the oracle model. The bottom row plots the Brier Scores. }
\end{figure}

@@

\section{Application to Real Data}
\label{Sec:real}

@@

\begin{table}[ht]
\centering
\caption{\label{tb:shedden} The C Index and Brier Score of all elastic net models selected by cross-validation methods (Section~\ref{Sec:real})}
\begin{tabular}{lccccccc}
\hline
  & \multicolumn{7}{c}{Cross-validated C Index} \\ 
  & $\alpha = 0.1$ & $\alpha = 0.2$ & $\alpha = 0.3$ & $\alpha = 0.5$ & $\alpha = 0.7$ & $\alpha = 0.9$ & $\alpha = 1$ \\ \cline{2-8}
V \& VH        & 0.645 &0.643 &0.630 &0.626 &0.617 &0.617 &0.617 \\
Basic          & 0.649 &0.650 &0.650 &0.649 &0.650 &0.651 &0.651 \\
Linear Pred    & 0.648 &0.650 &0.650 &0.649 &0.648 &0.617 &0.617 \\
Deviance Resid & 0.645 &0.647 &0.646 &0.626 &0.617 &0.617 &0.617 \\ 
\addlinespace
  & \multicolumn{7}{c}{Brier Score} \\ 
  & $\alpha = 0.1$ & $\alpha = 0.2$ & $\alpha = 0.3$ & $\alpha = 0.5$ & $\alpha = 0.7$ & $\alpha = 0.9$ & $\alpha = 1$ \\ \cline{2-8}
V \& VH        &0.231 &0.232 &0.243 &0.245 &0.249 &0.249 &0.249 \\
Basic          &0.226 &0.226 &0.225 &0.224 &0.221 &0.221 &0.221 \\
Linear Pred    &0.228 &0.226 &0.225 &0.224 &0.224 &0.249 &0.249 \\
Deviance Resid &0.231 &0.229 &0.228 &0.245 &0.249 &0.249 &0.249 \\ 
\hline
\end{tabular}
\end{table}

@@

\begin{figure}[!htb]
  \centering
  \includegraphics[height= 8cm ]{shedden.png}
  \caption{\label{fig:CVE} Rescaled cross-validated errors at $\alpha = 0.3$ (Section~\ref{Sec:real}). The horizonal axis is $\log(\lambda)$. The vertical axis is the cross-validated error rescaled to a range from 0 to 1 in order to allow plotting of the four methods on a common axis. The vertical dashed lines represent the $\lambda$ selected by a specific method. The linear predictor approach selected the smallest $\lambda$ and Verweij and Van Houwelingen's approach selected the greatest $\lambda$.}
\end{figure}

@@

\begin{table}[ht]
\centering
\caption{\label{Tab:go} Five most significantly enriched molecular functions among genes selected by the linear predictor method with the elastic net ($\alpha=0.3$) for the data described in Section~\ref{Sec:real}.  The five terms have a false discovery rate of 3\% using the Benjamini-Hochberg method. Enrichment: frequency of the annotation among the selected genes divided by its frequency in the genome as a whole.}
\begin{tabular}{llrrr}
\toprule
ID & Name & Selected & Enrichment & $p$-value \\
\midrule
GO:0008134 & transcription factor binding & 10 & 4.1 & 0.00015 \\
GO:0035257 & nuclear hormone receptor binding & 5 & 7.4 & 0.00057 \\
GO:0019869 & chloride channel inhibitor activity & 2 & 53.2 & 0.00061 \\
GO:0003700 & DNA-binding transcription factor activity & 14 & 2.6 & 0.00068 \\
GO:0005102 & signaling receptor binding & 16 & 2.4 & 0.00083 \\
\bottomrule
\end{tabular}
\end{table}

@@

Although we focused upon the penalized Cox regression with a lasso penalty in this work, the methods that we analyzed are applicable in any context that involves conducting model selection for survival models that are built upon the Cox partial likelihood, such as fitting penalized Cox regression with other penalty functions (elastic net \citep{Zou2005}, SCAD \citep{FanLi2001}, MCP \citep{Zhang2010}, precision lasso \citep{Wang2018}), other penalty structures (group selection, bi-level selection), and selecting functional forms of covariates in unpenalized Cox regression.
